---
title: "HW3"
author: "Eric Morris"
date: "10/5/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")
```


## Problem 1 

Loading in the BRFSS Data: 

```{r brfss_data_import}
library(p8105.datasets)

brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response_factor = as.factor(response)) %>% 
  rename(state = locationabbr, county = locationdesc)
```

In 2002, which states were observed at 7 locations?

```{r seven_locations}
brfss_data %>% 
  group_by(state) %>% 
  distinct(county) %>% 
  summarize(n_state = n()) %>% 
  filter(n_state == 7)
```

CT, MT, NH, NM, OR, TN, UT were observed at 7 locations.

Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r spaghetti}
brfss_data %>% 
  group_by(state, year) %>% 
  ggplot(aes(x = year, y = sample_size, color = state)) + 
  geom_line()
```

Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r tables}
brfss_data %>% 
  group_by(state, county, response, year, data_value) %>% 
  filter(year == 2002 | year == 2006 | year == 2010) %>% 
  filter(state == "NY") %>% 
  filter(response == "Excellent") %>% 
  summarize(mean_excellent = mean(data_value, na.rm = TRUE),
            sd_excellent = sd(data_value, na.rm = TRUE))
```


For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r avg_prop}

```

## Problem 2

